---
title: "PA1_template"
author: "Lidiia Sh"
output: html_document
---

##Loading and preprocessing the data

Load the data:
```{r}
activity<-read.csv("activity.csv", stringsAsFactors=FALSE)
```

General information about data:
```{r}
head(activity)
```
```{r}
str(activity)
```

Transform the data into a format suitable for analysis:
```{r}
activity$date<-as.Date(activity$date)
```

##What is mean total number of steps taken per day?

Calculate the total number of steps taken per day:
```{r}
Stepsperday<-aggregate(activity$steps,list(activity$date),sum)
```

Histogram of the total number of steps taken each day:
```{r}
hist(Stepsperday[,2],breaks=20,xlab="total number of steps taken per day",main="Histogram of total number of steps taken per day",col="green")
```

The mean and median of the total number of steps taken per day:
```{r}
n_mean<-mean(Stepsperday[,2],na.rm=TRUE)
n_median<-median(Stepsperday[,2],na.rm=TRUE)
```
Mean total number of steps taken per day is `r n_mean`.Median is `r n_median`.

##What is the average daily activity pattern?

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r}
forplot<-aggregate(activity$steps,list(activity$interval),mean,na.rm=TRUE)
plot(forplot,type="l",xlab="5-minute interval",ylab="Average number of steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxsteps<-forplot[which(forplot[,2]==max(forplot[,2]),arr.ind=TRUE),1]
```
The interval with the maxium number of steps across all days is `r maxsteps`.

##Imputing missing values

The total number of missing values in the dataset:
```{r}
sum(is.na(activity))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in. The mean for that 5-minute interval is a strategy for filling in all of the missing values in the dataset.
```{r}
activity_new<-activity
for (i in 1:dim(activity_new)[1]){
  if(is.na(activity_new[i,1])){
    activity_new[i,1]<-round(forplot[which(forplot[,1]==activity_new[i,3]),2])
  }
}
```

General information about new dataset:
```{r}
head(activity_new)
```
```{r}
str(activity_new)
```

Make a histogram of the total number of steps taken each day:
```{r}
Stepsperday_new<-aggregate(activity_new$steps,list(activity_new$date),sum)
hist(Stepsperday_new[,2],breaks=20,xlab="total number of steps taken per day",main="Histogram of total number of steps taken per day",col="blue")
```

Calculate the mean and median total number of steps taken per day:
```{r}
n_mean_new<-mean(Stepsperday_new[,2])
n_median_new<-median(Stepsperday_new[,2])
```
Mean total number of steps taken per day is `r n_mean_new`. Median is `r n_median_new`.

Do these values differ from the estimates from the first part of the assignment?
```{r}
n_mean-n_mean_new
```

```{r}
n_median-n_median_new
```
The imputing missing data on the estimates of the total daily number of steps leads to decrease the mean and median values.

##Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day:
```{r}
activity_new$weekday<-factor(weekdays(activity_new$date)==c("Sunday","Saturday"),levels=c(FALSE,TRUE),labels=c("weekday","weekend"))
head(activity_new)
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis):
```{r}
forplot_new<-aggregate(activity_new$steps,list(activity_new$interval,activity_new$weekday),mean)
library(lattice)
xyplot(forplot_new[,3]~forplot_new[,1] | forplot_new[,2],type="l",layout=c(1,2),xlab="Interval",ylab="Number of steps")
```